Stat 534 Project: \\ Extrapolation from Poisson Process Intensity Surface Models

### 2017-03-24 project/paper.rnw
Now the presentation is really done.

<<uxoexampledata>>=
# Define the window for the UXO site.
site_window <- owin(poly = cbind(x = c(1564294, 1564495, 1556870, 1557126),
                                 y = c(535421, 541130, 541085, 535576)),
                    unitname = c('foot', 'feet'))
@

@@

<<truth, cache = TRUE, fig.width = 4, fig.pos = 't', fig.cap = 'True intensity at the simulated UXO site.'>>=
# Bivariate normal intensity function that I originalally wrote for my
# writing project, parameterized in terms of the major and minor axes scales
# and the angle of the major axis from horizontal.
#   a is horizontal axis, b is vertical axis, r is rotation angle
gauss.elliptic <- function(x, y, mu.x = 0, mu.y = 0, s.a = 1, s.b = 1,
                           r = 0, maxrate = 1){
  rot <- zapsmall(matrix(c(cos(r), sin(r), -sin(r), cos(r)), nrow = 2))
  ab <- diag(c(s.a^2, s.b^2))
  sigma <- rot %*% ab %*% t(rot)
  siginv <- solve(sigma)
  mu <- matrix(c(mu.x, mu.y), nrow = 2)
  mat <- matrix(rbind(x, y), nrow = 2)
  return(maxrate * apply(mat, 2, function(vec){
      exp(-t(vec - mu) %*% siginv %*% (vec - mu) / 2)
    }))
}

@@

<<sim2000obs, fig.pos = 't', fig.cap = 'Observed geomagnetic anomalies from one realization of the UXO site. The transects are not shown on this plot because they would obscure the points.'>>=
# Load the sample data used in my writing project.
# The anomaly file contains the locations of the observed anamolies and the
# cog file contains the course-over-ground waypoints on the transects.
ex_anom <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.anomaly')
ex_path <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.cog')

@@

<<polyfit1, cache = TRUE, warning = FALSE, fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = ''>>=
# Create a quadrature scheme manually to ensure there are dummy points on
# each transect.
# TODO: Figure out weights.
ex_Q <- quad(data = ex_ppp,
             dummy = as.ppp(rbind(
               data.frame(vertices(Window(ex_ppp))),
               do.call(rbind, lapply(unique(ex_path$x), function(x){
                 return(data.frame(
                   x = x, y = seq(min(ex_path$y), max(ex_path$y),
                                  length.out = 128)
                 ))}))), W = Window(ex_ppp)))

@@

# Multiply predictions by 43560 to convert from anomalies per square foot
# to anomalies per acre.
plot(43560 * ex_p2$estimate, main = '2nd Degree')
plot(43560 * ex_p3$estimate, main = '3rd Degree')
plot(43560 * ex_p4$estimate, main = '4th Degree')
plot(43560 * ex_p5$estimate, main = '5th Degree')
plot(43560 * ex_p6$estimate, main = '6th Degree')
plot(43560 * ex_p7$estimate, main = '7th Degree')
plot(43560 * ex_p8$estimate, main = '8th Degree')
plot(43560 * ex_p9$estimate, main = '9th Degree')
plot(43560 * ex_p10$estimate, main = '10th Degree')
plot(43560 * ex_p11$estimate, main = '11th Degree')
plot(43560 * ex_p12$estimate, main = '12th Degree')
plot(43560 * density(ex_ppp_full) * area(ex_ppp_full) / ex_ppp_full$n,
     main = 'Kernel Density')
@

<<polyse1, fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = ''>>=
par(mfrow = c(4, 3), mar = c(1, 0, 2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree\nlog(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree\nlog(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree\nlog(SE)')
plot(ex_p2$se / ex_p2$estimate, main = '2nd Degree\nRSE')
plot(ex_p4$se / ex_p4$estimate, main = '4th Degree\nRSE')
plot(ex_p6$se / ex_p6$estimate, main = '6th Degree\nRSE')

>>>>

<<uxoexampledata, cache = TRUE>>=
# Define the window for the UXO site.
site_window <- owin(poly = cbind(x = c(1564294, 1564495, 1556870, 1557126),
                                 y = c(535421, 541130, 541085, 535576)),
                    unitname = c('foot', 'feet'))
@

@@

<<truth, cache = TRUE, dependson = 'uxoexampledata', fig.width = 4, fig.pos = 't', fig.cap = 'True intensity at the simulated UXO site.'>>=
# Bivariate normal intensity function that I originalally wrote for my
# writing project, parameterized in terms of the major and minor axes scales
# and the angle of the major axis from horizontal.
#   a is horizontal axis, b is vertical axis, r is rotation angle
gauss.elliptic <- function(x, y, mu.x = 0, mu.y = 0, s.a = 1, s.b = 1,
                           r = 0, maxrate = 1){
  rot <- zapsmall(matrix(c(cos(r), sin(r), -sin(r), cos(r)), nrow = 2))
  ab <- diag(c(s.a^2, s.b^2))
  sigma <- rot %*% ab %*% t(rot)
  siginv <- solve(sigma)
  mu <- matrix(c(mu.x, mu.y), nrow = 2)
  mat <- matrix(rbind(x, y), nrow = 2)
  return(maxrate * apply(mat, 2, function(vec){
      exp(-t(vec - mu) %*% siginv %*% (vec - mu) / 2)
    }))
}

@@

<<sim2000obs, cache = TRUE, dependson = 'uxoexampledata', fig.pos = 't', fig.cap = 'Observed geomagnetic anomalies from one realization of the UXO site. The transects are not shown on this plot because they would obscure the points.'>>=
# Load the sample data used in my writing project.
# The anomaly file contains the locations of the observed anamolies and the
# cog file contains the course-over-ground waypoints on the transects.
ex_anom <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.anomaly')
ex_path <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.cog')

@@

<<polyfit1, cache = TRUE, dependson = 'sim2000obs', warning = FALSE, fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = ''>>=
# Do a simple fit where spatstat sets up the quadrature so we can see where it
# puts the dummy points.
ex_fit1 <- ppm(ex_ppp ~ x + y)

# Create a quadrature scheme manually to ensure there are dummy points on
# each transect.
# TODO: Figure out weights.
ex_Q <- quad(data = ex_ppp, dummy = runifpoint(500000, Window(ex_ppp)))

@@

# Multiply predictions by 43560 to convert from anomalies per square foot
# to anomalies per acre.
plot(43560 * ex_p2$estimate, main = '2nd Degree')
plot(43560 * ex_p3$estimate, main = '3rd Degree')
plot(43560 * ex_p4$estimate, main = '4th Degree')
plot(43560 * ex_p5$estimate, main = '5th Degree')
plot(43560 * ex_p6$estimate, main = '6th Degree')
plot(43560 * ex_p7$estimate, main = '7th Degree')
plot(43560 * ex_p8$estimate, main = '8th Degree')
plot(43560 * ex_p9$estimate, main = '9th Degree')
plot(43560 * ex_p10$estimate, main = '10th Degree')
plot(43560 * ex_p11$estimate, main = '11th Degree')
plot(43560 * ex_p12$estimate, main = '12th Degree')
plot(43560 * density(ex_ppp_full) * area(ex_ppp_full) / area(ex_ppp),
     main = 'Kernel Density')
@

<<polyse1, cache = TRUE, dependson = 'polyfit1', fig.width = 6, fig.height = 8, fig.pos = 'p', fig.cap = ''>>=
par(mfrow = c(4, 3), mar = c(1, 0, 2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree\nlog(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree\nlog(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree\nlog(SE)')
plot(ex_p2$se / ex_p2$estimate, main = '2nd Degree\nRSE')
plot(ex_p4$se / ex_p4$estimate, main = '4th Degree\nRSE')
plot(ex_p6$se / ex_p6$estimate, main = '6th Degree\nRSE')

### 2017-03-24 project/pres.rnw
Now the presentation is really done.

<<uxoexampledata>>=
site_window <- owin(poly = cbind(x = c(1564294, 1564495, 1556870, 1557126),
                                 y = c(535421, 541130, 541085, 535576)),
                    unitname = c('foot', 'feet'))
ex_anom <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.anomaly')
ex_path <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.cog')
ex_win <- intersect.owin(site_window,
  do.call(union.owin, lapply(unique(ex_path$x), function(x){
    return(owin(c(x - 3, x + 3), c(535421, 541130),
           unitname = c('foot', 'feet')))
  })))
ex_ppp <- ppp(ex_anom$x, ex_anom$y, window = ex_win)
ex_ppp_full <- ppp(ex_anom$x, ex_anom$y, window = site_window)
@

@@

<<truth, cache = TRUE>>=
gauss.elliptic <- function(x, y, mu.x = 0, mu.y = 0, s.a = 1, s.b = 1,
                           r = 0, maxrate = 1){
  rot <- zapsmall(matrix(c(cos(r), sin(r), -sin(r), cos(r)), nrow = 2))
  ab <- diag(c(s.a^2, s.b^2))
  sigma <- rot %*% ab %*% t(rot)
  siginv <- solve(sigma)
  mu <- matrix(c(mu.x, mu.y), nrow = 2)
  mat <- matrix(rbind(x, y), nrow = 2)
  return(maxrate * apply(mat, 2, function(vec){
      exp(-t(vec - mu) %*% siginv %*% (vec - mu) / 2)
    }))
}
x <- seq(1556880, 1564495, by = 20)
y <- seq(535431, 541130, by = 20)
intense.mat <- matrix(100, nrow = length(x), length(y))
for(i in seq_along(x)){
  for(j in seq_along(y)){
    intense.mat[i, j] <- 100 + gauss.elliptic(x[i], y[j],
        mu.x = 1558400, mu.y = 540000,
        s.a = 800 / (2 * qnorm(0.995)), s.b = 1200 / (2 * qnorm(0.995)),
        r = pi/6, maxrate = 200
      ) + gauss.elliptic(x[i], y[j],
        mu.x = 1562000, mu.y = 537000,
        s.a = 2000 / (2 * qnorm(0.995)), s.b = 900 / (2 * qnorm(0.995)),
        r = 0, maxrate = 200
      )
  }
}
intense.im <- im(t(intense.mat), x, y, unitname = c('foot', 'feet'))

@@

<<polyfit, warning = FALSE, cache = TRUE>>=
ex_fit1 <- ppm(ex_ppp ~ x + y)
ex_Q <- quad(data = ex_ppp, dummy = runifpoint(500000, Window(ex_ppp)))

@@

<<resplots1>>=
par(mfrow = c(2, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)
plot(43560 * ex_p2$estimate, main = '2nd Degree Polynom')
plot(43560 * ex_p3$estimate, main = '3rd Degree Polynom')
plot(43560 * ex_p4$estimate, main = '4th Degree Polynom')
plot(43560 * ex_p5$estimate, main = '5th Degree Polynom')
plot(43560 * ex_p6$estimate, main = '6th Degree Polynom')
plot(43560 * ex_p7$estimate, main = '7th Degree Polynom')
@

@@

<<resplots2>>=
par(mfrow = c(2, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)
plot(43560 * ex_p8$estimate, main = '8th Degree Polynom')
plot(43560 * ex_p9$estimate, main = '9th Degree Polynom')
plot(43560 * ex_p10$estimate, main = '10th Degree Polynom')
plot(43560 * ex_p11$estimate, main = '11th Degree Polynom')
plot(43560 * ex_p12$estimate, main = '12th Degree Polynom')
plot(43560 * density(ex_ppp_full) * area(ex_ppp_full) / area(ex_ppp),
     main = 'Kernel Density')
@

@@

<<seplots>>=
par(mfrow = c(2, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree log(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree log(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree log(SE)')
plot(log(43560 * ex_p8$se), main = '8th Degree log(SE)')
plot(log(43560 * ex_p10$se), main = '10th Degree log(SE)')
plot(log(43560 * ex_p12$se), main = '12th Degree log(SE)')
@

@@

<<dummy>>=
par(mfrow = c(1, 2), mar = c(1, 0, 1, 0), cex = 1)
plot(ex_fit1$Q$dummy, pch = '.', border = NA,
     main = 'Default Dummy Points')
plot(site_window, add = TRUE)
plot(ex_Q$dummy, pch = '.', border = NA,
     main = 'Manual Dummy Points')
plot(site_window, add = TRUE)
@

@@

\item The magnitudes of the predictions are much too large

@@

\item What scale are the prediction SEs on?

>>>>

<<uxoexampledata, cache = TRUE>>=
site_window <- owin(poly = cbind(x = c(1564294, 1564495, 1556870, 1557126),
                                 y = c(535421, 541130, 541085, 535576)),
                    unitname = c('foot', 'feet'))
ex_anom <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.anomaly')
ex_path <- read.csv('easy_sample_tTA2_p200_bg100_fg200_rep2000.cog')
ex_win <- intersect.owin(site_window,
  do.call(union.owin, lapply(unique(ex_path$x), function(x){
    return(owin(c(x - 3, x + 3), c(535421, 541130),
           unitname = c('foot', 'feet')))
  })))
ex_ppp <- ppp(ex_anom$x, ex_anom$y, window = ex_win)
ex_ppp_full <- ppp(ex_anom$x, ex_anom$y, window = site_window)
@

@@

<<truth, cache = TRUE, dependson = 'uxoexampledata'>>=
gauss.elliptic <- function(x, y, mu.x = 0, mu.y = 0, s.a = 1, s.b = 1,
                           r = 0, maxrate = 1){
  rot <- zapsmall(matrix(c(cos(r), sin(r), -sin(r), cos(r)), nrow = 2))
  ab <- diag(c(s.a^2, s.b^2))
  sigma <- rot %*% ab %*% t(rot)
  siginv <- solve(sigma)
  mu <- matrix(c(mu.x, mu.y), nrow = 2)
  mat <- matrix(rbind(x, y), nrow = 2)
  return(maxrate * apply(mat, 2, function(vec){
      exp(-t(vec - mu) %*% siginv %*% (vec - mu) / 2)
    }))
}
x <- seq(1556880, 1564495, by = 20)
y <- seq(535431, 541130, by = 20)
intense.mat <- matrix(100, nrow = length(x), length(y))
for(i in seq_along(x)){
  for(j in seq_along(y)){
    intense.mat[i, j] <- 100 + gauss.elliptic(x[i], y[j],
        mu.x = 1558400, mu.y = 540000,
        s.a = 800 / (2 * qnorm(0.995)), s.b = 1200 / (2 * qnorm(0.995)),
        r = pi/6, maxrate = 200
      ) + gauss.elliptic(x[i], y[j],
        mu.x = 1562000, mu.y = 537000,
        s.a = 2000 / (2 * qnorm(0.995)), s.b = 900 / (2 * qnorm(0.995)),
        r = 0, maxrate = 200
      )
  }
}
intense.im <- im(t(intense.mat), x, y, unitname = c('foot', 'feet'))

@@

<<polyfit, warning = FALSE, cache = TRUE, dependson = 'uxoexampledata'>>=
ex_fit1 <- ppm(ex_ppp ~ x + y)
ex_Q <- quad(data = ex_ppp, dummy = runifpoint(500000, Window(ex_ppp)))

@@

<<resplots1, cache = TRUE, dependson = 'polyfit'>>=
par(mfrow = c(2, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)
plot(43560 * ex_p2$estimate, main = '2nd Degree Polynom')
plot(43560 * ex_p3$estimate, main = '3rd Degree Polynom')
plot(43560 * ex_p4$estimate, main = '4th Degree Polynom')
plot(43560 * ex_p5$estimate, main = '5th Degree Polynom')
plot(43560 * ex_p6$estimate, main = '6th Degree Polynom')
plot(43560 * ex_p7$estimate, main = '7th Degree Polynom')
@

@@

<<resplots2, cache = TRUE, dependson = 'polyfit'>>=
par(mfrow = c(2, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)
plot(43560 * ex_p8$estimate, main = '8th Degree Polynom')
plot(43560 * ex_p9$estimate, main = '9th Degree Polynom')
plot(43560 * ex_p10$estimate, main = '10th Degree Polynom')
plot(43560 * ex_p11$estimate, main = '11th Degree Polynom')
plot(43560 * ex_p12$estimate, main = '12th Degree Polynom')
plot(43560 * density(ex_ppp_full) * area(ex_ppp_full) / area(ex_ppp),
     main = 'Kernel Density')
@

@@

<<seplots, cache = TRUE, dependson = 'polyfit'>>=
par(mfrow = c(2, 3), mar = c(1, 0, 1.2, 3), las = 2, cex = 1)
plot(log(43560 * ex_p2$se), main = '2nd Degree log(SE)')
plot(log(43560 * ex_p4$se), main = '4th Degree log(SE)')
plot(log(43560 * ex_p6$se), main = '6th Degree log(SE)')
plot(log(43560 * ex_p8$se), main = '8th Degree log(SE)')
plot(log(43560 * ex_p10$se), main = '10th Degree log(SE)')
plot(log(43560 * ex_p12$se), main = '12th Degree log(SE)')
@

@@

<<dummy, cache = TRUE, dependson = 'polyfit'>>=
par(mfrow = c(1, 2), mar = c(1, 0, 1, 0), cex = 1)
plot(ex_fit1$Q$dummy, pch = '.', border = NA,
     main = 'Default Dummy Points')
plot(site_window, add = TRUE)
plot(ex_Q$dummy, pch = '.', border = NA,
     main = 'Manual Dummy Points')
plot(site_window, add = TRUE)
@

@@

@@
